04-analisis

### 2024-05-08 02-antecedentes.qmd
update plots and tables antecedentes y metodos

The question arises whether adults and children justify unequal access to welfare services based on market justice principles. Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013; @pierson_increasing_2000; @weaver_paths_2010], citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter. Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples. For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services. Figure 1 illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare. Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education. Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.

```{r}
#| fig-cap: "Average of market justice preferences by country"
```

@@

```{r}
#| fig-cap: "Summary of hypotheses"
```

>>>>

The question arises whether adults and children justify unequal access to welfare services based on market justice principles. Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013; @pierson_increasing_2000; @weaver_paths_2010], citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter. Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples. For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services. @fig-issp illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare. Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education. Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.

```{r echo=FALSE}
pacman::p_load(dplyr, ggplot2, haven)
load("input/data/issp_rec.RData")
```

```{r echo=FALSE}
#| label: fig-issp
#| fig-cap: "Average of market justice preferences by country"

base$country <- as.character(base$country)
base$country <- substr(base$country, 4, nchar(base$country))
base$country <- as_factor(base$country)

base %>% ggplot(aes(mean, country)) +
  theme_bw() +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  stat_summary(aes(xintercept = after_stat(x), y = 0), fun = mean, geom = "vline", orientation = "y", colour = "green") +
  xlab("") +
  ylab("Country")+
  scale_x_continuous(limits = c(1, 5),
                     breaks = seq(1, 5, by = 1),
                     label = c("Very unjust,\n definitely wrong", 
                               "Somewhat unjust,\n wrong", 
                               "Neither just nor\n unjust, mixed feelings", 
                               "Somewhat just,\n right", 
                               "Very just,\n definitely right")) +
  scale_color_discrete(name = NULL) +
  scale_shape_discrete(name = NULL) +
  theme(text = element_text(size = 9.35),
#        legend.position = "bottom",
        plot.margin = margin(b = 0))
```

@@

```{r echo=FALSE}
#| fig-cap: "Summary of hypotheses"
#| label: fig-hypotheses

knitr::include_graphics("input/img/hypotheses.png")
```

### 2024-05-08 03-methods.qmd
render paper
update plots and tables antecedentes y metodos

---
title: "03-methods"
author: "Equipo Edumer"
format: pdf
editor: visual
---

@@

Table 1 shows the items used, their response categories, and their frequencies.

```{r}
#| tbl-cap: "Dependent variables"
```

@@

The rest of the independent variables are divided into individual and school levels. At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home. Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection. Table 2 shows the items used, their response categories, and their frequency.

```{r}
#| tbl-cap: "Independent variables"
```

The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees. Table 3 shows the items used, response categories, and frequency.

```{r}
#| tbl-cap: "School context variables"
```

@@

The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open ScorrectlyOSF), the access to the document is available at this \[link\](https://doi.org/10.17605/OSF.IO/UFSDV). The statistical analysis of this research was performed using the free software R version 4.1.3.

>>>>

---
author: "Equipo Edumer"
format: pdf
editor: visual
---

@@

@tbl-desc-dependientes shows the items used, their response categories, and their frequencies.

```{r echo=FALSE}
pacman::p_load(summarytools, dplyr)
load("input/data/dependientes.RData")

```

```{r, results='asis', echo=FALSE}
#| tbl-cap: "Dependent variables"
#| label: tbl-desc-dependientes

df<-dfSummary(dependientes,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

df$Variable <- NULL # delete variable column

print(df)
```

@@

The rest of the independent variables are divided into individual and school levels. At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home. Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection. @tbl-desc-independent shows the items used, their response categories, and their frequency.

```{r echo=FALSE}
load("input/data/independientes.RData")
```

```{r, results='asis', echo=FALSE}
#| tbl-cap: "Independent variables"
#| label: tbl-desc-independent

df<-dfSummary(independientes,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

df$Variable <- NULL # delete variable column

print(df)
```

The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees. @tbl-desc-school shows the items used, response categories, and frequency.

```{r echo=FALSE}
load("input/data/school.RData")
```

```{r, results='asis', echo=FALSE}
#| tbl-cap: "School context variables"
#| label: tbl-desc-school

df<-dfSummary(school,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

df$Variable <- NULL # delete variable column

print(df)
```

@@

The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), the access to the document is available at this [link](https://doi.org/10.17605/OSF.IO/UFSDV). The statistical analysis of this research was performed using the free software R version 4.1.3.

### 2024-05-08 04-analisis.qmd
render paper
update plots and tables antecedentes y metodos

---
title: "04-analisis"
author: "Equipo Edumer"
format: pdf
editor: visual
---

@@

```{r}
#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
```

>>>>

---
author: "Equipo Edumer"
format: pdf
editor: visual
---

@@

```{r echo=FALSE}
load("input/data/deservingness.RData")
load("input/data/soc_effort.RData")
load("input/data/soc_talent.RData")
load("input/data/school_effort.RData")
load("input/data/school_talent.RData")
pacman::p_load(ggpubr)
```

```{r echo=FALSE}
#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
#| fig-asp: 1.5

plot_soc_effort <- soc_effort %>% ggplot(aes(esfuerzo_soc, mean)) +
  theme_bw(base_size = 10) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  xlab("Social effort") +
  ylab("Market justice preferences") +
  scale_y_continuous(limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_discrete(name = NULL) +
  scale_shape_discrete(name = NULL)

plot_deservingness <- deservingness %>% ggplot(aes(merito_soc, mean)) +
  theme_bw(base_size = 10) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  xlab("Deservingness") +
  ylab("Market justice preferences") +
  scale_y_continuous(limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_discrete(name = NULL) +
  scale_shape_discrete(name = NULL)

plot_soc_talent <- soc_talent %>% ggplot(aes(inteligencia_soc, mean)) +
  theme_bw(base_size = 10) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  xlab("Social talent") +
  ylab("Market justice preferences") +
  scale_y_continuous(limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_discrete(name = NULL) +
  scale_shape_discrete(name = NULL)

plot_school_effort <- school_effort %>% ggplot(aes(esfuerzo_esc, mean)) +
  theme_bw(base_size = 10) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  xlab("School effort") +
  ylab("") +
  scale_y_continuous(limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_discrete(name = NULL) +
  scale_shape_discrete(name = NULL)

plot_school_talent <- school_talent %>% ggplot(aes(inteligencia_esc, mean)) +
  theme_bw(base_size = 10) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  xlab("School talent") +
  ylab("") +
  scale_y_continuous(limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_discrete(name = NULL) +
  scale_shape_discrete(name = NULL)

ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent, plot_school_talent, plot_deservingness, ncol = 2, nrow = 3, common.legend = TRUE)
```
